---
title: "covid 19"
author: "Precious Calderon"
date: "Summer 2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

At first glance, it can be seen that there are more females alive than there are males, and there are more males dead than there are females. No concrete conclusion can be drawn from this information so far. Therefore, further testing must be conducted to determine whether dying of covid is independent of sex. 

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

`

```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")

```

The first bar chart shows the frequency of females and males that are alive or dead. It shows that there are generally more females in the sample, and among the female population, more than 1700 are alive while less than 100 females are dead from covid. For men, on the other hand, there are less than 1500 who are alive, and the number of dead males is larger than dead females. 

The second bar chart, meanwhile, shows a visual approximation of the percentages of females and males that are alive or dead. Among the female population, there seem to be a 60:40 ratio of living females to dead females. Among males, there seem to be a 40:60 ratio of living males to dead males. From the information gathered from the bar charts above, there appears to be more dead males than there are females. This statement will be tested further in the following sections of this write up. 


  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
The two tables above present information about the percentage of alive and dead females and males. Looking at the first table, it shows that among people who are alive, 56% of this population are females while 44% are male. Hence, more females are alive than males. Furthermore, among the people who are dead, 36% are female while 64% are male. In other words, there are more dead males than there are females. 

The second table, meanwhile, looks into the female and male population separately. The most vital information on this table is that there is only 1.4% dead among females while there is a 3% death among males. The death percentage among men is twice than the death rate of females. Therefore, it supports the idea that there is more deaths among male than females. Again, this statement will further be tested in the coming sections. 



## Inferential results 




## Conclusions

